Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-03-03 demystify.tex
updates

@@

\section{Can I negotiate my PhD offer?}\label{sec:negotiate}
For stipend, unlikely. TA is set by the university and while RA is set by the adviser, it is often based on the department's standard and it would be unfair to other students if you get more.  For start date or TA assignment (e.g., TA a specific course), it is uncommon but you can ask for it. Also, there is typically no moving allowance for PhD students. In short, things that are standard and set by the university or department are unlikely to be negotiable.  However, you can ask for things such as computer equipments (\autoref{sec:buying-equipment}), books, (and sometimes even furniture) and conference travel budget.

>>>>

\section{I am a member of an underrepresented group in CS}\label{sec:urm}
One question that often comes up is if you should mention, e.g., in SOP, that you are a member of an underrepresented and minority (URM) group. This question arises often for international (and even domestic) students. The main worry is that (i) they might get discriminated against (as often happen in their countries or personal experiences), and (ii) mentioning this might make them look like they are asking for special treatment.

In my personal opinion, the short answer is \emph{yes}, it can help, especially if you can articulate your experiences and how you can contribute to diversity and inclusion in CS. Most universities have a strong commitment to diversity and inclusion and have specific programs to \emph{increase the number of URM students} in their programs. Moreover, many universities and industrial organizations (e.g., Google) also have specific scholarships and fellowships for URM students (so you should look for specific scholarships for URM students).

It is completely understandable that some students might feel uncomfortable mentioning this in fear that they lost a chance of working with a prof who might be biased against URM students.  However, if a prof. has such an issue, then you might not want to work with them anyway (and it is likely that the university would not tolerate such behavior). In my experience, it is the opposite: many profs. are very supportive of URM students and many are actively looking for URM students to work with, and mentioning this in your application can make you stand out. It is also worth noting that professors often seek grants from government sponsors such as NSF, which often have broadening participation requirements.  However, if in doubt, you seek advice from your close professors and collaborators about this.

@@

\section{Can I negotiate my PhD offer?}\label{sec:negotiate}
For stipend, unlikely, as TA is set by the university and while RA is set by the adviser, it is often based on the department's standard and it would be unfair to other students if you get more. However, your stipend is often increased every year (by a small amount and automatically) and will be more after you pass your qualifying exam.

For start date or TA assignment (e.g., TA a specific course), it is uncommon but you can ask for it. Also, there is typically no moving allowance for PhD students. In short, things that are standard and set by the university or department are unlikely to be negotiable.  However, you can ask for things such as computer equipments (\autoref{sec:buying-equipment}), books, (and sometimes even furniture) and conference travel budget.
